A comprehensive approach to analyzing environmental data with non-detects

### 2024-05-24 Rmarkdown/paper.Rmd
generated figures

<!-- to do: -->

<!-- 1. add a linear model for site to the Ti GAM (currently there in only alpha) -->

@@

The data and code necessary to reproduce the main results from the paper are available at https://github.com/bentrueman/ADD-REPOSITORY-NAME-HERE; several functions used to fit the models in Stan are available in a separate R package [@bgamcar1]. We used R version 4.3.3 throughout [@Rbase], along with a collection of contributed packages [@tidyverse; @brms; @cmdstanr; @assertr; @here; @patchwork; @PNWColors].

@@

```{r figure-1, fig.cap="**Figure 1. (a)** One iteration of the linear regression simulation. The model that accounts for left-censoring via the cumulative distribution function recovered the true model parameters well, whereas the naive model that used substitution at one-half the detection limit was biased. **(b)** The same pattern was evident across the entire simulation: the censored regression model recovered the true parameters well and the naive model was biased."}
knitr::include_graphics(here("R/figures/figure-simulation.png"))
```

@@

```{r figure-2, fig.cap="**Figure 2.** Titanium concentration time series representing biosolids collected at three locations (light lines). Model predictions are superimposed in bold, the shaded regions represent a 95% credible interval on the posterior mean, and non-detects are shown as vertical dashed lines extending to the detection limit. A single value beyond the plot limits is annotated."}
knitr::include_graphics(here("R/figures/figure-censored-response.png"))
```

@@

```{r figure-3, fig.cap="**Figure 3.** Cobalt concentrations as a function of cadmium concentrations in biosolids from three treatment facilities. Nondetect cadmium concentrations are represented as horizontal lines extending to the detection limit. A robust (Student t) linear model is superimposed in blue and the shaded region represents a 95% credible interval on the posterior mean. The equivalent non-robust model yields an extremely wide credible interval due to the unusually high cobalt concentration of 309 µg g^-1^. Coordinates outside the plotting limits are annotated in parentheses."}
knitr::include_graphics(here("R/figures/figure-censored-predictor.png"))
```

@@

```{r figure-4, fig.cap="**Figure 4. (a)** Pairwise Bayesian correlations among the elemental concentrations in the dataset, estimated using Student t and Gaussian likelihoods. **(b)** The robust model---fitted with Student t likelihoods---identifies more correlation than the non-robust model fitted with Gaussian likelihoods."}
knitr::include_graphics(here("R/figures/figure-correlation-matrix.png"))
```

@@

```{r figure-5, fig.cap="**Figure 5. (a)** The dataset projected onto the first two probabilistic principal components (*z* in equation (6)). Values appearing outside the extents of the plot are annotated in parentheses at the margins. **(b)** The first two principal axes; that is, the orthonormalized columns of the transformation matrix $W$."}
knitr::include_graphics(here("R/figures/figure-ppca.png"))
```

>>>>

@@

The data and code necessary to reproduce the main results from the paper are available at https://github.com/bentrueman/censored-env-data-analysis; several functions used to fit the models in Stan are available in a separate R package [@bgamcar1]. We used R version 4.3.3 throughout [@Rbase], along with a collection of contributed packages [@tidyverse; @brms; @cmdstanr; @assertr; @here; @patchwork; @PNWColors].

@@

```{r figure-1, fig.cap="**Figure 1. (a)** One iteration of the linear regression simulation. The model that accounts for left-censoring via the cumulative distribution function recovered the true model parameters well, whereas the naive model that used substitution at one-half the detection limit was biased. **(b)** The same pattern was evident across the entire simulation: the censored regression model recovered the true parameters well and the naive model was biased."}
knitr::include_graphics(here("figures/figure-simulation.png"))
```

@@

```{r figure-2, fig.cap="**Figure 2.** Titanium concentration time series representing biosolids collected at three locations (light lines). Model predictions are superimposed in bold, the shaded regions represent a 95% credible interval on the posterior mean, and non-detects are shown as vertical dashed lines extending to the detection limit. A single value beyond the plot limits is annotated."}
knitr::include_graphics(here("figures/figure-censored-response.png"))
```

@@

```{r figure-3, fig.cap="**Figure 3.** Cobalt concentrations as a function of cadmium concentrations in biosolids from three treatment facilities. Nondetect cadmium concentrations are represented as horizontal lines extending to the detection limit. A robust (Student t) linear model is superimposed in blue and the shaded region represents a 95% credible interval on the posterior mean. The equivalent non-robust model yields an extremely wide credible interval due to the unusually high cobalt concentration of 309 µg g^-1^. Coordinates outside the plotting limits are annotated in parentheses."}
knitr::include_graphics(here("figures/figure-censored-predictor.png"))
```

@@

```{r figure-4, fig.cap="**Figure 4. (a)** Pairwise Bayesian correlations among the elemental concentrations in the dataset, estimated using Student t and Gaussian likelihoods. **(b)** The robust model---fitted with Student t likelihoods---identifies more correlation than the non-robust model fitted with Gaussian likelihoods."}
knitr::include_graphics(here("figures/figure-correlation-matrix.png"))
```

@@

```{r figure-5, fig.cap="**Figure 5. (a)** The dataset projected onto the first two probabilistic principal components (*z* in equation (6)). Values appearing outside the extents of the plot are annotated in parentheses at the margins. **(b)** The first two principal axes; that is, the orthonormalized columns of the transformation matrix $W$."}
knitr::include_graphics(here("figures/figure-ppca.png"))
```
